\bf Hierarchical Network Structure Promotes Dynamical Robustness

### 2015-07-27 tex/overview.tex
update explanation of relative contribution of SCCs to robustness

When maximizing robustness subject to a fixed number $N$ of nodes, we can make use of our previous result in \ref{eq:sccmaxrobustness} to restrict attention to directed graphs with the maximal number of links between SCCs.  Hence, we can replace $l$ with $l_{\mathrm{max}}$ in \ref{eq:robtype} to obtain
\begin{equation}
1-R(G) = \frac{\sum_{i=1}^n k_i d_i (1-R_i)}
  {\frac{1}{2} N^2 - \frac{1}{2} \sum_{i=1}^n k_i v_i^2 + \sum_{i=1}^n k_i d_i} .
\end{equation}
The condition on the total number of nodes means that the $k_i$'s cannot be varied independently but are subject to the constraint $\sum_{i=1}^n k_i v_i = N$.  Geometrically, this constraint defines a hyperplane and the interesection of this hyperplane with the positive orthant is the simplex whose vertices are the points $(N/v_1, 0, \ldots, 0), (0, N/v_2, \ldots, 0), \ldots, (0, 0, \ldots, N/v_n)$.

Since both numerator and denominator are linear in the variables $k_i$ over which we are optimizing, an elementary argument shows the quantity will assume its minimum value at one of the vertices of the simplex.
The value at the $i$-th vertex equals
\begin{equation}\label{eq:robqual}
 q_i =
 \frac{2 d_i (1 - R_i)}
      {(k_i - 1) v_i^2 + 2 d_i},
\end{equation}
hence the network that maximizes robustness will have as many SCCs as possible be of the type which have the smallest value of $q_i$. For the SCC consisting of a single node that has a link connecting to itself $\lim_{N \rightarrow \infty} Nq = 1$. For other SCCs of two and three nodes this quantity is computed in \refsupp{} \ref{tab:structstabmat} and \ref{tab:structstabmat3}.

@@

>>>>

When maximizing robustness subject to a fixed number $N$ of nodes, we can make use of our previous result in \ref{eq:sccmaxrobustness} to restrict attention to directed graphs with the maximal number of links between SCCs.  Hence, we can replace $l$ with $l_{\mathrm{max}}$ in \ref{eq:robtype} to obtain
\begin{equation}
1-R(G) = \frac{\sum_{i=1}^n k_i d_i (1-R_i)}
  {\frac{1}{2} N^2 - \frac{1}{2} \sum_{i=1}^n k_i v_i^2 + \sum_{i=1}^n k_i d_i} .
\end{equation}
The condition on the total number of nodes means that the $k_i$'s cannot be varied independently but are subject to the constraint $\sum_{i=1}^n k_i v_i = N$.  Geometrically, this constraint defines a hyperplane and the interesection of this hyperplane with the positive orthant is the simplex whose vertices are the points $(N/v_1, 0, \ldots, 0), (0, N/v_2, \ldots, 0), \ldots, (0, 0, \ldots, N/v_n)$. When $v_i = 1$ this corresponds to the dynamical robustness of the SCC with a single node and link to itself.

Since both numerator and denominator are linear in the variables $k_i$ over which we are optimizing, the quantity will assume its minimum value at one of the vertices of the simplex.
The value at the $i$-th vertex equals
\begin{equation}\label{eq:robqual}
 q_i =
 \frac{2 d_i (1 - R_i)}
      {(k_i - 1) v_i^2 + 2 d_i},
\end{equation}
hence the network that maximizes robustness will have as many SCCs as possible be of the type which have the smallest value of $q_i$.

% $\lim_{N \rightarrow \infty} Nq = 1$

@@

For the SCC consisting of a single node that has a link connecting to itself, which is the only SCC in the network corresponding to the total ordering on the network variables, \ref{eq:robqual} gives $q= \frac{1}{N+1}$. Assuming this value needs to be smaller than the corresponding value of $q_i$ for any other SCC, $\frac{1}{N+1} < q_i$, and simplifying the algebra, results in a sufficient condition for networks that contain only the single node SCC to have a higher value of dynamical robustness than networks containing any combination of alternative SCCs:
\begin{equation}\label{eq:maxrobcond}
\frac{2 d_i}{v_i} (1-R_i) > 1.
\end{equation}
For SCCs of two and three nodes this quantity is computed in \refsupp{} \ref{tab:structstabmat} and \ref{tab:structstabmat3}, where all of these SCCs satisfy this condition. On this basis, we conjecture that, in addition to the fact that the most dynamically robust network architectures correspond to the total ordering on a given collection of SCCs demonstrated by the analysis leading up to \ref{eq:robschematic}, that the most dynamically robust network architectures also correspond to the graph associated to the total ordering on all network variables.

### 2015-07-27 tex/table2x2stab.tex
update explanation of relative contribution of SCCs to robustness

\begin{table*}[h]
\begin{center}
\begin{tabular}{ c || c | c | c | c | c }
\hline
matrix & \specialcell{number\\of SCCs} & connectivity & \specialcell{dynamical \\ robustness} & \specialcell{probability\\of stability} & \specialcell{q \\ (\ref{eq:robqual})}\\
\hline
  $\begin{pmatrix}
a & b \\
d & c
\end{pmatrix}$ & 1 & 4 & 0.62 & 0.25 & 1.52\\
  $\begin{pmatrix}
a & b \\
d & 0
\end{pmatrix}$, $\begin{pmatrix}
0 & b \\
d & c
\end{pmatrix}$ & 1 & 3 & 0.5 & 0.25 & 1.5\\
  $\begin{pmatrix}
a & 0 \\
d & c
\end{pmatrix}$, $\begin{pmatrix}
a & b \\
0 & c
\end{pmatrix}$ & 2 & 3 & 0.67 & 0.25 & \\
$\begin{pmatrix}
a & 0 \\
0 & c
\end{pmatrix}$ & 2 & 2 & 0.5 & 0.25 & \\
\end{tabular}
\end{center}
\caption{{\bf Dynamical robustness and probability of stability for two variable systems derived analytically}. All matrices not listed have $0$ probability of stability.}\label{tab:structstabmat}
\end{table*}
% \twocolumn

>>>>

\begin{table*}[h]
\begin{center}
\begin{tabular}{ c || c | c | c | c | c }
\hline
\specialcell{adjacency \\ matrix} & \specialcell{number\\of SCCs} & connectivity & \specialcell{dynamical \\ robustness} & \specialcell{probability\\of stability} & \specialcell{$\frac{2d_i}{v_i}(1-R_i)$ \\ (\ref{eq:maxrobcond})}\\
\hline
  $\begin{pmatrix}
1 & 1 \\
1 & 1
\end{pmatrix}$ & 1 & 4 & 0.62 & 0.25 & 1.52\\
  $\begin{pmatrix}
1 & 1 \\
1 & 0
\end{pmatrix}$, $\begin{pmatrix}
0 & 1 \\
1 & 1
\end{pmatrix}$ & 1 & 3 & 0.5 & 0.25 & 1.5\\
  $\begin{pmatrix}
1 & 0 \\
1 & 1
\end{pmatrix}$, $\begin{pmatrix}
1 & 1 \\
0 & 1
\end{pmatrix}$ & 2 & 3 & 0.67 & 0.25 & \\
$\begin{pmatrix}
1 & 0 \\
0 & 1
\end{pmatrix}$ & 2 & 2 & 0.5 & 0.25 & \\
\end{tabular}
\end{center}
\caption{{\bf Dynamical robustness and probability of stability for two variable systems derived analytically}. All matrices not listed have $0$ probability of stability.}\label{tab:structstabmat}
\end{table*}
% \twocolumn
